ML-based approach for accelerating global search algorithm for solving multicriteria problems\thanks{This study was supported by the Russian Science Foundation, project No 21-11-00204.}

### 2022-02-14 2022/LION/main.tex
submitted

The paper reflects results of a new research direction connected with utilizing ML methods for acceleration of the global search algorithm in the case of its application to MCO problems. The proposed approach is based on building a number of separation planes in the criteria space to segregate the Pareto set and accelerate the process of its construction. The efficiency of the approach is estimated experimentally compared to several other MCO methods.

@@

In the course of Pareto set assessment all trial points are partitioned into two classes. This bipartition is based on belonging to the Pareto set. For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}. Examples of separating planes are shown in Fig.~\ref{fig:1}. During the choice of a new trial point, the distances of all obtained criteria values from the separating plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.

@@

In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$)
\sout{machine learning} \textcolor[rgb]{1,0,0}{use of separating planes in the criteria space}. 
Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly. Other algorithm's parameters were the same. The averaged results can be found in Table \ref{tab:2}.

>>>>

The paper reflects results of a new research direction connected with utilizing ML methods for acceleration of the global search algorithm in the case of its application to MCO problems. The proposed approach is based on building a number of separating planes in the criteria space to segregate the Pareto set and accelerate the process of its construction. The efficiency of the approach is estimated experimentally compared to several other MCO methods.

@@

In the course of Pareto set assessment all trial points are partitioned into two classes. This bipartition is based on belonging to the Pareto set. For these classes, a separating hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}. Examples of separating planes are shown in Fig.~\ref{fig:1}. During the choice of a new trial point, the distances of all obtained criteria values from the separating plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.

@@

In order to estimate efficiency of machine learning procedure in ML\_MGSA, this method tested in 2 variants: with ($\alpha>0$) and without ($\alpha=0$)
use of separating planes in the criteria space. 
Each multicriterial problem was reduced to 50 scalar problems (\ref{eq:4}) with different weight coefficients $\lambda$, distributed uniformly. Other algorithm's parameters were the same. The averaged results can be found in Table \ref{tab:2}.
